Journey: A Study of Scalabilty of Memory Management in Linux

### 2017-04-29 paper.tex
Addressed a couple more todos and added abstract

large memory systems

study scalability of linux kernel as first step

point to some results

@@

The memory overhead of \texttt{vm\_area\_struct}s increases linearly with the
number of mapped regions when these regions are separated by unused memory
addresses. When a workload maps a large number of regions, this overhead can be
of the order of hundreds of megabytes to gigabytes for large memory systems.  To
reduce this overhead, it is crucial that regions with similar access permissions
are placed adjacently in the address space so that they can be identified by a
single structure. It is important for large applications such as object-oriented
databases that map a large number of regions (TODO: Cite Linux book. Should we
mention this?)to be aware of these overheads and mitigate them.

@@

>>>>

Recent advances in memory technologies are making it possible for machines with
    large amounts of memory to become commonplace.  Much research is currently
    being done to study how non-volatile and large memory systems would change
    system design.  We see the need for a lightweight tool to study such
    systems at scale with real workloads.  The first step towards building such
    a tool is to study the scalability of existing memory management systems.

We observe that in most cases, Linux kernel memory management scales well with
    large amounts of physical memory. However, we observe that some workloads
    can trigger poor behavior. In particular, workloads that allocated many
    virtual memory areas can accumulate significant memory and computational
    overhead of kernel metadata.

We also observe that transparent hugepages can induce significantly earlier
    onset of memory pressure in workloads that have random access patterns,
    which may nullify the benefits from better TLB behavior. This can happen
    even in workloads that would not otherwise experience memory pressure.

Finally, we observe that the 99.99-th percentile of disk latency during swapping
can account for over 70\% of runtime for the workloads we tested.

@@

The memory overhead of \texttt{vm\_area\_struct}s increases linearly with the
number of mapped regions when these regions are separated by unused memory
addresses. When a workload maps a large number of regions, this overhead can be
of the order of hundreds of megabytes to gigabytes for large memory systems.  To
reduce this overhead, it is crucial that regions with similar access permissions
are placed adjacently in the address space so that they can be identified by a
single structure. It is important for large applications such as object-oriented
databases that map a large number of regions to be aware of these overheads and mitigate them \cite{utlk}.

@@

% TODO: we need to discuss implications for large memory in the insights sections more
